---
title: "Analysis of long data"
author: "Dominik Klepl"
date: "23 5 2018"
output: html_document
---

# Used libraries
```{r library}
pacman::p_load(rethinking,ggplot2,ggthemes,bayesplot)
```

# Load data
```{r load data}
data=read.csv("data/clean_data_long.csv")

#rename group_ID to just ID
colnames(data)[c(1,6)]=c("ID","trial")
data$ID = as.factor(data$ID)
```


Hypothesis 1: Friends tend to link more words together + interaction = friends' linked_n decreases with rounds whereas strangers' remain stable.
  models: friends ~ linked_n
          friends ~ linked_n * round

Hypothesis 2: Friends make more mistakes - performance is lower.
  models: friends ~ performance
          friends ~ performance * round 
            - It should improve with time for friends, there's not much to improve for                 strangers.
            
Hypothesis 3: Friends tend to overestimate their confidence in success
  models: friends ~ confidence 
          friends ~ confidence * round
            - should get better over time
          friends ~ confidence * is_success
            - when success (1) - confidence is higher than strangers'
            - when not success (0) - confidence is also higher than strangers'


# 1. Explore the data
```{r explore data}
psych::describeBy(data[,c(-1,-4:-6,-8)],data$friends)

#create condition variable just for plotting purposes
data$Condition=ifelse(data$friends==1,yes="Friends",no="Strangers")
data$Condition=as.factor(data$Condition)

```

Hypothesis 1
```{r}
#linked_n 
(explore_linked_n=ggplot(data,aes(x=Condition,y=linked_n,fill=Condition))+
  geom_boxplot()+
  labs(title="Number of linked words by condition",x="",y="number of linked words"))

#linked_n * round
(explore_linked_n_round=ggplot(data,aes(x=trial,y=linked_n,color=Condition))+
  geom_smooth(method = "glm")+
  labs(title="Average number of linked words over played rounds by condition",y="number of linked words"))

ggsave("Figures/Long/explore_linked.jpg",explore_linked_n)
ggsave("Figures/Long/explore_linked_n_round.jpg",explore_linked_n_round)
```

Hypothesis 2
```{r}
(explore_performance = ggplot(data,aes(x=Condition,y=performance,fill=Condition))+
  geom_boxplot()+
   labs(title="Performance by condition",x=""))

(explore_performance_round = ggplot(data,aes(x=trial,y=performance,color=Condition))+
  geom_smooth(method="glm")+
  labs(title="Average performance over played rounds by condition",y="performance"))

ggsave("Figures/Long/explore_performance.jpg",explore_performance)
ggsave("Figures/Long/explore_performance_round.jpg",explore_performance_round)
```

#TODO:Hypothesis 3 - only for SocKult paper
```{r}

```

# 2. Build models to test hypotheses
We build separate models for each hypotheses and then one that combines all the predictors.

## 2.1 Hypothesis 1 - Linked_n
```{r}
m_linked_n = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a + bL*linked_n ,
        a ~ dnorm(0,0.05),
        bL ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

#check the chains
plot(m_linked_n)
pairs(m_linked_n)

m_linked_n_random = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a[ID] + bL*linked_n ,
        a[ID] ~ dnorm(0,0.05),
        bL ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_linked_n_random)
pairs(m_linked_n_random)

m_linked_n_round = map2stan(
    alist(
      friends ~ dbinom( 1 , p ),
      logit(p) <- a +bL*linked_n +bT*trial +bLT*linked_n*trial,
      a ~ dnorm(0,0.05),
      bL ~ dnorm(0,0.2),
      bT ~ dnorm(0,0.2),
      bLT ~ dnorm(0,0.2)
  ),
    data=data,iter = 1e4,chains=2,cores = 2)

plot(m_linked_n_round)
pairs(m_linked_n_round)

#needs longer chains - low n_eff
m_linked_n_round_random = map2stan(
    alist(
      friends ~ dbinom( 1 , p ),
      logit(p) <- a[ID] +bL*linked_n +bT*trial +bLT*linked_n*trial,
      a[ID] ~ dnorm(0,0.1),
      bL ~ dnorm(0,0.2),
      bT ~ dnorm(0,0.2),
      bLT ~ dnorm(0,0.2)
  ),
    data=data,iter = 3e4,chains=2,cores = 2)

plot(m_linked_n_round_random)
pairs(m_linked_n_round_random)

compare(m_linked_n,m_linked_n_random,m_linked_n_round,m_linked_n_round_random)
```


## 2.2 Hypothesis 2 - Performance
```{r}
m_performance = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a + bP*performance ,
        a ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_performance)
pairs(m_performance)

m_performance_random = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a[ID] + bP*performance ,
        a[ID] ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_performance_random)
pairs(m_performance_random)

m_performance_round = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a + bP*performance + bT*trial + bPT*performance*trial ,
        a ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2),
        bT ~ dnorm(0,0.2),
        bPT ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_performance_round)
pairs(m_performance_round)

m_performance_round_random = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a[ID] + bP*performance + bT*trial + bPT*performance*trial ,
        a[ID] ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2),
        bT ~ dnorm(0,0.2),
        bPT ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_performance_round_random)
pairs(m_performance_round_random)

compare(m_performance,m_performance_random,m_performance_round,m_performance_round_random)
```


## TODO: 2.3 Hypothesis 3 - Confidence

## 2.4 Combinations of models and full model
```{r}
m_performance_linked_n= map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a[ID] + bP*performance +  bL*linked_n,
        a[ID] ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2),
        bL ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

m_full = map2stan(
    alist(
        friends ~ dbinom( 1 , p ) ,
        logit(p) <- a[ID] + bP*performance +  bL*linked_n + bT*trial + bPT*performance*trial + bLT*linked_n*trial,
        a[ID] ~ dnorm(0,0.05),
        bP ~ dnorm(0,0.2),
        bL ~ dnorm(0,0.2),
        bT ~ dnorm(0,0.2),
        bPT ~ dnorm(0,0.2),
        bLT ~ dnorm(0,0.2)
    ) ,
    data=data,
    iter = 1e4,chains = 2,cores=2)

plot(m_full)
pairs(m_full)
```


# 3. Compare models and assess model quality (PPC plots)
```{r}
comparison=compare(m_linked_n,m_linked_n_random,m_linked_n_round,m_linked_n_round_random,m_performance,m_performance_random,m_performance_round,m_performance_round_random,m_performance_linked_n,m_full)
comparison=round(comparison@output,3)

write.csv(comparison,"Results/Long/model_selection.csv")

#PPC plots - how could are the models predicting proportion of friend groups in the data

#define function that calculates the proportion
success <- function(x) sum(x == 1)/length(x)

#proportion in training data
y=data$friends
success(y)

#predictions of models
y_link_round=sim(m_linked_n_round_random)
y_performance_round=sim(m_performance_round_random)
y_performance_link=sim(m_performance_linked_n)
y_full = sim(m_full,n=1e4)

# + an ensemble model
ens = ensemble(m_linked_n_round_random,m_performance_round_random,m_performance_linked_n,m_full)

y_ens = ens$sim

ppc_stat(y,y_link_round,stat = 'success',binwidth = 0.01)
ppc_stat(y,y_performance_round,stat = 'success',binwidth = 0.01)
ppc_stat(y,y_performance_link,stat = 'success',binwidth = 0.01)
(PPC_full=ppc_stat(y,y_full,stat = 'success',binwidth = 0.01)+
  labs(title="Predicted proportion of group of friends and the real proportion",x="Proportion of friends"))
ppc_stat(y,y_ens,stat = 'success',binwidth = 0.01)

ggsave("Figures/Long/PPC_full.jpg",PPC_full)
```


# 4. Results of each effect ( use only best model(s) ) and plot it
```{r}
#m_full has the most Akaike's weight so we'll interpret the results from that model

(results_full = precis(m_full,depth = 2))
results_full = round(results_full@output,3)
plot(precis(m_full,depth = 2))

write.csv(results_full,"Results/Long/Results_m_full.csv")
```

Plot effect of performance
```{r}
#create counterfactual data - everything constant except for performance
perf=seq(0,100,by=5)

performance_vis = data.frame(
  performance=perf,
  linked_n = rep(0,length(perf)),
  trial = rep(0,length(perf)),
  friends = rep(0,length(perf)),
  ID= rep(sample(1:10,1),length(perf))
)

#compute the mean prediction + 89% intervals
perf_l = link(m_full,data=performance_vis,n=1e4)
performance_vis$mean=apply(perf_l, 2 , mean )
performance_vis=cbind(performance_vis,as.data.frame(t(apply(perf_l , 2 , PI , prob=0.89 ))))

#and plot
(performance_plot=ggplot(performance_vis,aes(x=performance))+
  geom_line(aes(y=mean))+
  geom_ribbon(aes(ymin=`5%`,ymax=`94%`),alpha=0.5,fill="#0092ff")+
  labs(title="Decrease of odds of being friends with increasing performance",x="performance in %",y="Odds"))

ggsave("Figures/Long/Performance_visualization.jpg",performance_plot)
```

Now repeat the same for linked_n
```{r}
#create counterfactual data - everything constant except for linked_n
linked=seq(0,15,by=1)

linked_n_vis = data.frame(
  performance=rep(0,length(linked)),
  linked_n = linked,
  trial = rep(0,length(linked)),
  friends = rep(0,length(linked)),
  ID= rep(sample(1:10,1),length(linked))
)

#compute the mean prediction + 89% intervals
linked_l = link(m_full,data=linked_n_vis,n=1e4)
linked_n_vis$mean=apply(linked_l, 2 , mean )
linked_n_vis=cbind(linked_n_vis,as.data.frame(t(apply(linked_l , 2 , PI , prob=0.89 ))))

#and plot
(linked_n_plot=ggplot(linked_n_vis,aes(x=linked_n))+
  geom_line(aes(y=mean))+
  geom_ribbon(aes(ymin=`5%`,ymax=`94%`),alpha=0.5,fill="#0092ff")+
  labs(title="Increase of odds of being friends with increasing number of linked words",x="n of linked words",y="Odds"))

ggsave("Figures/Long/Linked_n_visualization.jpg",linked_n_plot)
```

Compare prior and posterior
```{r}
samples_full = extract.samples(m_full,n=1e4)

prior=rnorm(0,0.2,n=1e4)

PP_performance = data.frame(Prior=prior,Posterior=samples_full$bP)
PP_linked = data.frame(Prior=prior,Posterior=samples_full$bL)

PP_performance = reshape2::melt(PP_performance)
PP_linked = reshape2::melt(PP_linked)

(PP_perf_plot=ggplot(PP_performance,aes(x=value,color=variable))+
  geom_line(stat = "density")+
  labs(title="Prior and posterior of the effect of performance"))

(PP_linked_plot=ggplot(PP_linked,aes(x=value,color=variable))+
  geom_line(stat = "density")+
  labs(title="Prior and posterior of the effect of linked_n"))

ggsave("Figures/Long/PP_performance.jpg",PP_perf_plot)
ggsave("Figures/Long/PP_linked_n.jpg",PP_linked_plot)

#what is the probability the linked_n effect is only positive (no overlap with 0)
post_linked_n = samples_full$bL
sum(post_linked_n>0)/length(post_linked_n)
```


